A Derivative Tracking Model for Wind Power Forecast Error

### 2020-07-23 LaTeX_Code/Paper/aoas-template.tex
Minor changes
Paper submitted!

\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{plots/wind_and_solar/plot_over_years.eps}
\caption{Installed wind and solar power over the years \cite{sultana2017review}. We recall the importance of accurate forecasts to use green energies optimally.}
\end{figure}

Reliable wind power generation forecasting is crucial for the following applications (see, for example, \cite[5]{gieb}, \cite[162]{chang}, \cite{zhbo}):
\begin{itemize}
\item Allocation of energy reserves such as water levels in dams or oil, and gas reserves.
\item Operation scheduling of controllable power plants.
\item Optimization of the price of electricity for different parties such as electric utilities, Transmission system operator (TSOs), Electricity service providers (ESPs), Independent power producers (IPPs), and energy traders.
\item Maintenance planning such as that of power plants components and transmission lines.

@@

\begin{figure}[H]
\centering
  \includegraphics[width=0.485\textwidth]{plots/31_2.eps}\quad\includegraphics[width=0.485\textwidth]{plots/31.eps}
  \caption{Pointwise confidence bands fitted, for the same daily segment, through diffusion models without derivative tracking (plot on the left) and with derivative tracking (plot on the right).}
  \label{fig:derivative_tracking}
\end{figure}

@@

It is worth mentioning that Jacobi diffusions have been successfully applied in several disciplines, among them finance (see \cite{vago} and references therein) and neuroscience (\cite{dotala}).

@@

\begin{equation}
\theta_t\geq \max\left(\frac{\alpha\theta_0+\dot p_t}{1-p_t},\frac{\alpha\theta_0-\dot p_t}{p_t}\right)\tag{B}. 
\end{equation} \label{condB}
The proof of this theoretical statement is presented in Section \ref{Appendix}. \\

Remark: Condition (B) shows that the time-varying parameter $\theta_t$ becomes unbounded when $p_t = 0$ or $p_t = 1$. Therefore, we consider the following truncated prediction function
\begin{equation}
  p_t^{\epsilon} = \left\{
  \begin{array}{@{}cl@{}}
    \epsilon \!\!\!&\text{if \quad} p_t  < \epsilon  \\
    p_t  \!\!\!&\text{if \quad}  \epsilon \leq p_t < 1 - \epsilon \\
    1 - \epsilon  \!\!\!&\text{if \quad}  p_t \geq 1 - \epsilon
 \end{array}\right.  \label{corrforecast}
\end{equation}
that satisfies $p_t^{\epsilon} \in [\epsilon, 1 - \epsilon]$ for any $0 < \epsilon < \frac{1}{2}$ and $t \in [0,T]$, providing that $\theta_t$ is bounded for every $t \in [0,T]$. \\

@@

\item {Quadratic variation:} We approximate the quadratic variation of the  It\^{o}'s  process $V$
\begin{equation*}
\langle V\rangle _t=\int_0^t b(V_s; \bm{\theta}, p_s)^2 \dif s, \text{ where }  \:\: b(V_s; \bm{\theta}, p_s) = 
\sqrt{2 \alpha \theta_0 (V_s +p_s ) (1-V_s - p_s)},
\end{equation*}
with the discrete sum $\sum_{0< t_{j, i-1} \leq t}\left(V_{t_{j, i}} - V_{t_{j, i-1}}\right)^2$.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/LL/Log-Likelihood.eps}\quad\includegraphics[width=0.485\textwidth]{plots/LL/Log-Likelihood_testing.eps}
\caption{{Negative log-likelihood's level sets for the training sets (plot on the left), and for the test sets (plot on the right). All optimal values are located over the curve $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively.}}
\label{fig:neg-LL}
\end{figure}
{On Figure (\ref{fig:neg-LL}), we can see the level sets for the negative log-likelihood for both training and test sets.} The numerical values of each relevant point can be seen in Table (\ref{tab:optimal_values}). We set the optimal parameters in the $V$-space $(\theta_0^V,\alpha^V)=(1.93,0.050)$, as it is where the negative log-likelihood for the training sets reaches its minimum value.
\begin{table}[H]
\centering
\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{3}{c}{Training sets} & \multicolumn{3}{c}{Test sets} \\
 \midrule
 & $\theta_0$ & $\alpha$ & $\theta_0\alpha$ & $\theta_0$ & $\alpha$ & $\theta_0\alpha$\\
 \midrule
 Initial guess & 1.54 & 0.072 & 0.111 & 1.96 & 0.053 & 0.104 \\
 fminsearch & 1.14 & 0.076 & 0.097 & 1.64 & 0.054 & 0.089 \\
 fmincon & 1.58 & 0.062 & 0.097 & 1.63 & 0.055 & 0.089 \\
 fminunc & 1.54 & 0.063 & 0.097 & 1.96 & 0.045 & 0.089 \\
 Evaluations & 1.93 & 0.050 & 0.097 & 1.59 & 0.056 & 0.089 \\
 \bottomrule
\end{tabular}
\caption{Value points from Figure \ref{fig:neg-LL}.}
\label{tab:optimal_values}
\end{table}
{We observe that all the local (possibly global) minimizers are located over the curves $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively. This effect shows that the optimization is more sensitive to the diffusion than to the drift.}\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Support information (funding), if any,   %%
%% should be provided in the                %%
%% Acknowledgements section.                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
This work was partially supported by the KAUST Office of Sponsored Research (OSR) under Award number URF/1/2584 -- 01 -- 01 in the KAUST Competitive Research Grants Program Round 8 and the Alexander von Humboldt Foundation.

>>>>

\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{plots/wind_and_solar/plot_over_years_IRENA.eps}
\caption{Worldwide installed wind and solar power 2010--2019 \citep{irena20}. We recall the importance of accurate forecasts to use green energies optimally.}
\end{figure}

%Reliable wind power generation forecasting is crucial for the following applications (see, for example, \cite[5]{gieb}, \cite[162]{chang}, \cite{zhbo}):
Reliable wind power generation forecasting is crucial for the following applications (see, for example, \cite{gieb}, \cite{chang}, \cite{zhbo}):
\begin{itemize}
\item Allocation of energy reserves such as water levels in dams or oil, and gas reserves.
\item Operation scheduling of controllable power plants.
\item Optimization of the price of electricity for different parties such as electric utilities, Transmission system operator (TSOs), Electricity service providers (ESPs), Independent power producers (IPPs), and energy traders.
\item Maintenance planning such as that of power plants components and transmission lines.

@@

\begin{figure}[H]
\centering
  \includegraphics[width=0.485\textwidth]{plots/1.eps}\quad\includegraphics[width=0.485\textwidth]{plots/bands/1.eps}
  \caption{Pointwise confidence bands fitted, for the same daily segment, through diffusion models without derivative tracking (plot on the left) and with derivative tracking (plot on the right).}
  \label{fig:derivative_tracking}
\end{figure}

@@

It is worth mentioning that Jacobi diffusions have been successfully applied in several disciplines, among them finance (see (\cite{vago}) and references therein) and neuroscience (\cite{dotala}).

@@

\begin{equation}
\theta_t\geq \max\left(\frac{\alpha\theta_0+\dot p_t}{1-p_t},\frac{\alpha\theta_0-\dot p_t}{p_t}\right)\tag{B}. 
\end{equation} \label{condB}
The proof of this theoretical statement is presented in the Appendix. \\

Remark: Condition (B) shows that the time-varying parameter $\theta_t$ becomes unbounded when $p_t = 0$ or $p_t = 1$. Therefore, we consider the following truncated prediction function
\begin{equation}
p_t^{\epsilon}=\begin{cases}
\epsilon&\quad\text{if}\quad p_t<\epsilon\\
p_t&\quad\text{if}\quad\epsilon\leq p_t<1-\epsilon\\
1-\epsilon&\quad\text{if}\quad p_t\geq1-\epsilon
\end{cases}
\label{corrforecast}
\end{equation}
that satisfies $p_t^{\epsilon} \in [\epsilon, 1 - \epsilon]$ for any $0 < \epsilon < \frac{1}{2}$ and $t \in [0,T]$, providing that $\theta_t$ is bounded for every $t \in [0,T]$. \\

@@

\item {Quadratic variation:} We approximate the quadratic variation of the  It\^{o}'s  process $V$
\begin{equation*}
\langle V\rangle _t=\int_0^t b(V_s; \bm{\theta}, p_s)^2 \dif s, \text{ where }  \:\: b(V_s; \bm{\theta}, p_s) = 
\sqrt{2 \alpha \theta_0 (V_s +p_s ) (1-V_s - p_s)},
\end{equation*}
with the discrete sum $\sum_{0< t_{j, i-1} \leq t}\left(V_{t_{j, i}} - V_{t_{j, i-1}}\right)^2$.\smallskip

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/LL/Log-Likelihood.eps}\quad\includegraphics[width=0.485\textwidth]{plots/LL/Log-Likelihood_testing.eps}
\caption{{Negative log-likelihood's level sets for the training sets (plot on the left), and for the test sets (plot on the right). All optimal values are located over the curve $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively.}}
\label{fig:neg-LL}
\end{figure}
{On Figure (\ref{fig:neg-LL}), we can see the level sets for the negative log-likelihood for both training and test sets.} The numerical values of each relevant point can be seen in Table (\ref{tab:optimal_values}). We set the optimal parameters in the $V$-space $(\theta_0^V,\alpha^V)=(1.93,0.050)$, as it is where the negative log-likelihood for the training sets reaches its minimum value.
\begin{table}[H]
\centering
\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{3}{c}{Training sets} & \multicolumn{3}{c}{Test sets} \\
 \midrule
 & $\theta_0$ & $\alpha$ & $\theta_0\alpha$ & $\theta_0$ & $\alpha$ & $\theta_0\alpha$\\
 \midrule
 Initial guess & 1.54 & 0.072 & 0.111 & 1.96 & 0.053 & 0.104 \\
 fminsearch & 1.14 & 0.076 & 0.097 & 1.64 & 0.054 & 0.089 \\
 fmincon & 1.58 & 0.062 & 0.097 & 1.63 & 0.055 & 0.089 \\
 fminunc & 1.54 & 0.063 & 0.097 & 1.96 & 0.045 & 0.089 \\
 Evaluations & 1.93 & 0.050 & 0.097 & 1.59 & 0.056 & 0.089 \\
 \bottomrule
\end{tabular}
\caption{Coordinates of the initial guess point and of the optimal points represented in Figure \ref{fig:neg-LL}.}
\label{tab:optimal_values}
\end{table}
{We observe that all the local (possibly global) minimizers are located over the curves $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively. This effect shows that the optimization is more sensitive to the diffusion than to the drift.}\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Support information (funding), if any,   %%
%% should be provided in the                %%
%% Acknowledgements section.                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
This research was partially supported by the KAUST Office of Sponsored Research (OSR) under Award number URF/1/2584 -- 01 -- 01 in the KAUST Competitive Research Grants Program Round 8, the Alexander von Humboldt Foundation, the chair Risques Financiers, Fondation du Risque, and the Laboratory of Excellence MME-DII Grant no. ANR11-LBX--0023--01 (\url{http://labex-mme-dii.u-cergy.fr/}). We thank UTE (\url{https://portal.ute.com.uy/}) for providing the data used in this research.
